---
title: "Data Preparation AirBnB United Kingdom"
authors: "Team 2: Sanne van Ettinger, Claudia Berkhof, Demi van de Pol, Jurg Jacobs, Rob van der Wielen"
date: "3/8/2022"
output: html_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	include = FALSE)
```

```{r, include=FALSE}
# Install packages ONLY ONCE!
#install.packages("plyr")
#install.packages("dplyr")
#install.packages("tidyverse")
#install.packages("ggplot2")
#install.packages("readr")
#install.packages("googledrive")
#install.packages("caret")
#install.packages("data.table")
#install.packages("R.utils")
#install.packages("dint")
#install.packages("zoo")
#install.packages("summarytools")
```

```{r, include=FALSE}
# Loading packages
library(plyr)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(readr)
library(googledrive)
library(caret)
library(data.table)
library(R.utils)
library(dint)
library(zoo)
library(summarytools)
```

```{r, include=FALSE}
# Import raw data
data_airbnb_uk <- data.frame(fread("../../data/data_airbnb_uk.csv"))
```

## Data Exploration & Preparation
## Content
1. Data Exploration
1.1 Explore data using summary statistics
1.2 Detect the origin of missing values

2. Data Preparation
2.1 Create new variables as reference to the quarterly data
2.2 Delete unnecessary columns
2.3 Rename variables "neighbourhood_cleansed" and "calculated_host_listings_count"
2.4 Rescale the variable "review_scores_rating"
2.5 Change the variable "price" to numeric
2.6 Removing NA's from the dataset
2.7 Detect outliers in the data


## 1. Data Exploration
### 1.1 Explore data using summary statistics

```{r}
head(data_airbnb_uk)
```

```{r}
summary(data_airbnb_uk)
```

#### Summary statistics variable: city
```{r}
dfSummary(data_airbnb_uk$city)
```

#### Summary statistics variable: neighbourhood_cleansed
```{r}
# Number of neighbourhoods in total
unique(data_airbnb_uk$neighbourhood_cleansed)
```

```{r}
# Count combinations of cities and neighbourhoods
data_airbnb_uk %>% 
  count(neighbourhood_cleansed, city, sort = TRUE)
```

#### Summary statistics variable: room_type
```{r}
dfSummary(data_airbnb_uk$room_type)
```

```{r}
# Plot room_type overall
ggplot(data_airbnb_uk, aes(room_type)) +
  geom_bar(stat = "count", fill = "red") +
  stat_count(geom = "text", colour = "black", size = 6,
             aes(label = ..count..), position = position_stack(vjust = 0.5)) +
  labs(y = "room_type_count")
```

```{r}
# Plot room_type per city
ggplot(aes(x = room_type), data = data_airbnb_uk) +
  geom_bar(aes(fill = city), colour = "grey20", lwd = 0.5) +
  stat_count(geom = "text", colour = "white", size = 5,
           aes(label=..count.., group=city), position = position_stack(vjust = 0.5)) +
  labs(y = "room_type_count")
```

#### Summary statistics London: neighbourhood_cleansed & room_type
```{r}
data_airbnb_london <- data_airbnb_uk %>% 
  filter(city == "london")
dfSummary(data_airbnb_london$neighbourhood_cleansed)
dfSummary(data_airbnb_london$room_type)
```

#### Summary statistics Great Manchester: neighbourhood_cleansed & room_type
```{r}
data_airbnb_great_manchester <- data_airbnb_uk %>% 
  filter(city == "great manchester")
dfSummary(data_airbnb_great_manchester$neighbourhood_cleansed)
dfSummary(data_airbnb_great_manchester$room_type)
```

#### Summary statistics Bristol: neighbourhood_cleansed & room_type
```{r}
data_airbnb_bristol <- data_airbnb_uk %>% 
  filter(city == "bristol")
dfSummary(data_airbnb_bristol$neighbourhood_cleansed)
dfSummary(data_airbnb_bristol$room_type)
```


### 1.2 Detect the origin of missing values

```{r}
# Create new dataset including only the observations that have NA's for review_scores_rating
data_airbnb_uk_na <- data_airbnb_uk[is.na(data_airbnb_uk$review_scores_rating),]
```

#### Detect origin missing values using host_is_superhost
```{r}
data_airbnb_uk %>%
  count(host_is_superhost)
```

```{r}
data_airbnb_uk_na %>%
  count(host_is_superhost)
```

#### Detect origin missing values using host_since
```{r}
data_host <- data_airbnb_uk %>%
  count(host_since)
```

```{r}
data_host_na <- data_airbnb_uk_na %>%
  count(host_since)
```

```{r}
# Plot the data for the dataset without NA's using ggplot
ggplot(data = data_host, aes(x = host_since, y = n)) +
  geom_bar(stat = "identity", fill = "purple") +
  labs(title = "Date of becoming a host",
              x = "Date", y = "Number of host")
```

```{r}
# Plot the data for the dataset with NA's using ggplot
ggplot(data = data_host_na, aes(x = host_since, y = n)) +
  geom_bar(stat = "identity", fill = "purple") +
  labs(title = "Date of becoming a host",
              x = "Date", y = "Number of host")
```

```{r}
# Convert data to years for the dataset without NA's
dates <- as.POSIXct(data_host$host_since, format="%Y-%m-%d")
dates_2 <- format(dates, format="%Y")
dates_2 <- as.data.frame(dates_2)

data_host$year <- dates_2$dates
```

```{r}
# Plot the converted data for the dataset without NA's using ggplot
ggplot(data = data_host, aes(x = year, y = n)) +
  geom_bar(stat = "identity", fill = "purple") +
  labs(title = "Date of becoming a host in years",
              x = "year", y = "Number of host")
```

```{r}
# Convert data to years for dataset with NA's
dates_na <- as.POSIXct(data_host_na$host_since, format="%Y-%m-%d")
dates_2_na <- format(dates_na, format="%Y")
dates_2_na <- as.data.frame(dates_2_na)

data_host_na$year <- dates_2_na$dates
```

```{r}
# Plot the converted data for the dataset with NA's using ggplot
ggplot(data = data_host_na, aes(x = year, y = n)) +
  geom_bar(stat = "identity", fill = "purple") +
  labs(title = "Date of becoming a host in years for the na dataset",
              x = "year", y = "Number of host")
```



## 2. Data Preparation
### 2.1 Create new variables as reference to the quarterly data

```{r}
# Create new variables: quarter, year, year_quarter
data_airbnb_uk$quarter <- get_quarter(data_airbnb_uk$last_scraped)
data_airbnb_uk$year <- get_year(data_airbnb_uk$last_scraped)
data_airbnb_uk$year_quarter <- as.yearqtr(data_airbnb_uk$last_scraped, format="%Y-%m-%d")
data_airbnb_uk$year_quarter <- format(data_airbnb_uk$year_quarter, format = "%Y-%q")
```


### 2.2 Delete unnecessary columns

```{r}
data_airbnb_uk_cleaned <- data_airbnb_uk %>% 
  select(id, city, quarter, year, year_quarter, neighbourhood_cleansed, room_type, calculated_host_listings_count, review_scores_rating, price)
```


### 2.3 Rename variables "neighbourhood_cleansed" and "calculated_host_listings_count"

```{r}
data_airbnb_uk_cleaned <- data_airbnb_uk_cleaned %>% 
  rename(neighbourhood_name = neighbourhood_cleansed, num_host_listings = calculated_host_listings_count)
```


### 2.4 Rescale the variable "review_scores_rating"

```{r}
data_airbnb_uk_cleaned <- data_airbnb_uk_cleaned %>% 
  mutate(review_scores_rating_rescaled = ifelse(review_scores_rating>5,review_scores_rating/20, review_scores_rating))
```


### 2.5 Change the variable "price" to numeric

```{r}
# Remove the dollar sign and commas (for prices above thousand)
data_airbnb_uk_cleaned$price <- gsub('[$,]', '', data_airbnb_uk_cleaned$price)

# Changing price variable from character to numeric
data_airbnb_uk_cleaned$price <- as.numeric(data_airbnb_uk_cleaned$price)

# Check for the results
class(data_airbnb_uk_cleaned$price)
```


### 2.6 Remove NA's from the dataset

```{r}
# Removing the missing values and create a new dataset without the NA's
data_airbnb_uk_cleaned <- data_airbnb_uk_cleaned %>% 
  drop_na(review_scores_rating, review_scores_rating_rescaled)
```


### 2.7 Detect outliers in the data (data_airbnb_uk_cleaned)
#### Detecting outliers num_host_listings 
```{r}
# Plot num_host_listings
boxplot(data_airbnb_uk_cleaned$num_host_listings,
  ylab = "num_host_listings_count")
```

```{r}
# Filter num_host_listings on more than 600 listings
data_airbnb_uk_outliers_I <- data_airbnb_uk_cleaned %>% 
  filter(num_host_listings > 600)
```

#### Detecting outliers: review_rating_score_rescaled
```{r}
# plot review_scores_rating_rescaled
boxplot(data_airbnb_uk_cleaned$review_scores_rating_rescaled,
  ylab = "review_scores_rating_rescaled_count")
```

#### Detecting outliers: price
```{r}
boxplot(data_airbnb_uk_cleaned$price,
  ylab = "price")
```

```{r}
data_airbnb_uk_outliers_II <- data_airbnb_uk_cleaned %>% 
  filter(price > 10000)
```



## 3. After cleaning
### 3.1 New summary statistics

#### Summary statistics on the complete dataset
```{r}
head(data_airbnb_uk_cleaned)
```

```{r}
summary(data_airbnb_uk_cleaned)
```

#### Summary statistics variable: city
```{r}
dfSummary(data_airbnb_uk_cleaned$city)
```

#### Summary statistics variable: room_type
```{r}
dfSummary(data_airbnb_uk_cleaned$room_type)
```

#### Summary statistics variable: room_type & city
```{r}
# Plot room_type per city
ggplot(aes(x = room_type), data = data_airbnb_uk_cleaned) +
  geom_bar(aes(fill = city), colour = "grey20", lwd = 0.5) +
  stat_count(geom = "text", colour = "white", size = 5,
           aes(label=..count.., group=city), position = position_stack(vjust = 0.5)) +
  labs(y = "room_type_count")
```

#### Summary statistics variable: neighbourhood_name 
```{r}
unique(data_airbnb_uk_cleaned$neighbourhood_name)
```

#### Summary statistics variable: quarter
```{r}
dfSummary(data_airbnb_uk_cleaned$quarter)
```



### 3.2 Write the cleaned data to a csv file

```{r}
# Exporting the cleaned data to a CSV file
dir.create("../../gen/analysis/input/")
fwrite(data_airbnb_uk_cleaned, '../../gen/analysis/input/data_airbnb_uk_cleaned.csv', row.names = F)
```














