---
title: "Data Preparation AirBnB United Kingdom"
authors: "Team 2: Sanne van Ettinger, Claudia Berkhof, Demi van de Pol, Jurg Jacobs, Rob van der Wielen"
date: "3/8/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	include = FALSE)
```

```{r Installing packages, include=FALSE}
# Install packages
install.packages("plyr")
install.packages("dplyr")
install.packages("tidyverse")
install.packages("ggplot2")
install.packages("readr")
install.packages("googledrive")
install.packages("caret")
install.packages("data.table")
install.packages("R.utils")
install.packages("dint")
install.packages("zoo")
```

```{r Loading packages, include=FALSE}
# Loading packages
library(plyr)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(readr)
library(googledrive)
library(caret)
library(data.table)
library(R.utils)
library(dint)
library(zoo)
```

```{r Import raw data, include=FALSE}
# Import raw data
setwd("../../data")
data_airbnb_uk <- read.csv("data_airbnb_uk.csv")
```


________
The raw data has been prepared and cleaned, using a variety of common data operations in R, involving dplyr, tidyverse, or data.table. Common operations are merging, aggregating, de-deduplication, reshaping, converting dates, or using regular expressions. Basic programming concepts are made use of appropriately to increase speed and minimize errors (e.g., looping, vectorization, writing functions, handling errors/debugging). Additional variables are created from the raw data (feature engineering).
________

## Introduction
This RMarkdown displays the process of cleaning and preparing the data for analysis. 

## Content
1. Create new variables as reference to the quarterly data
2. Delete unnecessary columns
3. Rescale the variable "review_scores_rating"
4. Change the variable "price" to numeric
5. Removing NA's from the dataset

## 1. Create new variables as reference to the quarterly data
As the data of 2021 is collected per quarter, and the purpose of this research is to compare different time periods, several new variables named "quarter", "year" and "year_quarter" should be created. By doing this, a distinction can be made between the different periods in time. These new variables are stored in a new dataset: data_airbnb_uk_cleaned. 

```{r Create new variables: quarter, year, year_quarter}
# creating new variables
data_airbnb_uk$quarter <- get_quarter(data_airbnb_uk$last_scraped)
data_airbnb_uk$year <- get_year(data_airbnb_uk$last_scraped)
data_airbnb_uk$year_quarter <- as.yearqtr(data_airbnb_uk$last_scraped, format="%Y-%m-%d")
data_airbnb_uk$year_quarter <- format(data_airbnb_uk$year_quarter, format = "%Y-%q")
```

## 2. Delete unnecessary columns
As this research focuses on the main question "Which types of rooms in which neighborhoods benefited the most from the Corona Pandemic in the United Kingdom?", some variables in the original dataset might not be needed to come to an answer. For that reason only the following variables are kept in the dataset to maintain a clear overview of the variables needed for this research:
* id: Airbnb's unique identifier for the listing.
* city: The city in which the Airbnb is located. 
* quarter: The quarter in which the Airbnb is listed.
* year: The year in which the Airbnb is listed. 
* year_quarter: The year and quarter in which in the Airbnb is listed.  
* neighbourhood_cleansed: The neighbourhood in which the Airbnb is located. 
* room_type: The type of room the Airbnb offers. 
* calculated_host_listings_count: The number of listings the host has in the current scrape, in the city / region geography. 
* host_listings_count: The number of listings the host has (per Airbnb calculations).
* review_scores_rating: The overall rating of the Airbnb based on the review scores on a 0 to 5 scale.
* price: The price at which the Airbnb is listed.


```{r Delete unnecessary columns}
data_airbnb_uk_cleaned <- data_airbnb_uk %>% 
  select(id, city, quarter, year, year_quarter, neighbourhood_cleansed, room_type, calculated_host_listings_count, review_scores_rating, price)
```

The variable names of "neighbourhood_cleansed" and "calculated_host_listings_count" are not revealing much of what they stand for. Therefore, these two variables need to be renamed so that their names are informative by themselves: 

```{r Rename variables neighborhood_cleansed and calculated_host_listings_count}
data_airbnb_uk_cleaned %>% 
  rename(neighbourhood_name = neighbourhood_cleansed, num_host_listings = calculated_host_listings_count)
```

## 3. Rescale the variable "review_scores_rating"
During the data exploration it was observed that the range of the variable "review_scores_rating" was a bit out of proportion.By inspecting the data it becomes clear that the range of this variables has changed over time. In quarter 1 of 2021 the variable "review_scores_rating" ranges between 0 and 100. However, in the following quarters (2, 3 and 4) the variable "review_scores_rating" ranges between 0 and 5. This will lead to problems while analyzing the data. Therefore this needs to be taken care off by rescaling the variable: 

```{r Rescale the variable review_scores_rating}
data_airbnb_uk_cleaned <- data_airbnb_uk_cleaned %>% 
  mutate(review_scores_rating_rescaled = ifelse(review_scores_rating>5,review_scores_rating/20, review_scores_rating))
```

## 4. Change the variable "price" to numeric
From the data exploration became clear that the variable "price" in this data is imported as a character. For further analyses the variable "price" needs to be changed into a numeric. Before this can be done the dollar sign and the comma for prices above thousand have to be removed. 

```{r Change the variable price}
# Remove the dollar sign and commas (for prices above thousand)
data_airbnb_uk_cleaned$price <- gsub('[$,]', '', data_airbnb_uk_cleaned$price)

# Changing price variable from character to numeric
data_airbnb_uk_cleaned$price <- as.numeric(data_airbnb_uk_cleaned$price)

# Check for the results
class(data_airbnb_uk_cleaned$price)
```

## 5. Remove NA's from the dataset
In the data exploration attention was paid to NA's (= missing values) in the data, specific for the variable review_rating_score, by means of plots and summary statistics. From the plots and summary statistics it remains unclear what the direct cause is of the missing values in the dataset. However, it is expected that the missing values of the variable review_rating_score are caused by guests or hosts that do not write a review. As the observations with the missing values can not be used for further analyses, they should be removed from the dataset. 

```{r Removing the missing values}
# Removing the missing values and create a new dataset without the NA's
data_airbnb_uk_cleaned <- data_airbnb_uk_cleaned %>% 
  drop_na(review_scores_rating, review_scores_rating_rescaled)
```


```{r Exporting the cleaned data to a CSV file}
# Exporting the cleaned data to a CSV file
write.csv(data_airbnb_uk_cleaned, '../../data/data_airbnb_uk_cleaned.csv', row.names = F)
```














